---
title: "Final Project"
output: charan kumar chaganabaoina
date: "2024-11-27"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Loading necessary libraries
library(dplyr)
library(ggplot2)
```

```{r}
# Loading data
airbnb_data <- read.csv("C:/Users/saich/Documents/New folder charan/AB_NYC_2019.csv")

# 1. Checking missing values
colSums(is.na(airbnb_data))
```

```{r}
# 2. Handling missing values
airbnb_data$reviews_per_month[is.na(airbnb_data$reviews_per_month)] <- 0
```

```{r}
colSums(is.na(airbnb_data))
```
```{r}
# Checking the structure of the dataset to see data types of each column
str(airbnb_data)

```

```{r}

# 3. Converting 'last_review' to Date
airbnb_data$last_review <- as.Date(airbnb_data$last_review)
```

```{r}
head(airbnb_data)
```

```{r}
# 1. Summary of key columns
summary(airbnb_data$price)  # Summary statistics for price
summary(airbnb_data$number_of_reviews)  # Summary for number of reviews
```

```{r}
#Mean, Median, Standard Deviation of Price
mean_price <- mean(airbnb_data$price, na.rm = TRUE)  # Mean
median_price <- median(airbnb_data$price, na.rm = TRUE)  # Median
sd_price <- sd(airbnb_data$price, na.rm = TRUE)  # Standard Deviation

#Mean, Median, Standard Deviation of Number of Reviews
mean_reviews <- mean(airbnb_data$number_of_reviews, na.rm = TRUE)  # Mean
median_reviews <- median(airbnb_data$number_of_reviews, na.rm = TRUE)  # Median
sd_reviews <- sd(airbnb_data$number_of_reviews, na.rm = TRUE)  # Standard Deviation

# Printing the results
cat("Price Summary:\n")
cat("Mean:", mean_price, "Median:", median_price, "SD:", sd_price, "\n")

cat("Number of Reviews Summary:\n")
cat("Mean:", mean_reviews, "Median:", median_reviews, "SD:", sd_reviews, "\n")
```
```{r}
# Creating the box plot
ggplot(airbnb_data, aes(x = neighbourhood_group, y = price, fill = neighbourhood_group)) +
  geom_boxplot(outlier.colour = "red", outlier.size = 2) +  # Highlight outliers
  scale_y_continuous(limits = c(0, 500)) +  # Set price limit for better visualization
  labs(
    title = "Price Variation Across Neighbourhood Groups",
    x = "Neighbourhood Group",
    y = "Price"
  ) +
  theme_minimal()

```
###The box plot shows how Airbnb prices differ in various parts of New York City. Manhattan stands out with the highest average prices and the widest range of prices. In contrast, Bronx, Queens, and Staten Island tend to have lower and more consistent prices. All areas, however, have some listings that are much more expensive than usual, which are shown as outliers on the plot.

```{r}
# Bar plot for room type distribution
ggplot(airbnb_data, aes(x = neighbourhood_group, fill = room_type)) +
  geom_bar(position = "dodge") +
  labs(
    title = "Distribution of Room Types by Neighbourhood Group",
    x = "Neighbourhood Group",
    y = "Count",
    fill = "Room Type"
  ) +
  theme_minimal()

```
##This bar chart shows the different types of Airbnb rooms available in each neighbourhood group. Manhattan and Brooklyn have the highest number of listings, especially for entire homes/apartments and private rooms. Shared rooms are the least common across all neighbourhoods.

```{r}
# Map of listings
ggplot(airbnb_data, aes(x = longitude, y = latitude, color = neighbourhood_group)) +
  geom_point(alpha = 0.5, size = 0.7) +  # Add scatter points with some transparency
  labs(
    title = "Map of Airbnb Listings in NYC",
    x = "Longitude",
    y = "Latitude",
    color = "Neighbourhood Group"
  ) +
  theme_minimal()

```
##This map shows where Airbnb listings are located in New York City. Each dot represents a listing, and the colors show different neighborhoods. Most listings are in Manhattan, Brooklyn, and Queens, while fewer are in the Bronx and Staten Island.

```{r}
# Calculating revenue potential
airbnb_data$revenue_potential <- airbnb_data$price * airbnb_data$availability_365

# View the first few rows to check the new column
head(airbnb_data)

```
```{r}
# Summary statistics for revenue potential
summary(airbnb_data$revenue_potential)

```
```{r}
# Boxplot for Revenue Potential by Neighbourhood Group
ggplot(airbnb_data, aes(x = neighbourhood_group, y = revenue_potential, fill = neighbourhood_group)) +
  geom_boxplot(outlier.colour = "red", outlier.size = 2) +
  scale_y_continuous(limits = c(0, 500000)) +  # Adjust limits for better visualization
  labs(
    title = "Revenue Potential by Neighbourhood Group",
    x = "Neighbourhood Group",
    y = "Revenue Potential"
  ) +
  theme_minimal()

```
##This box plot shows the revenue potential of Airbnb listings in different New York City neighborhoods. Manhattan and Brooklyn have the highest potential earnings, while the Bronx, Queens, and Staten Island show lower earnings. There are some listings in all neighborhoods with very high revenue potential, shown as outliers.


```{r}
# Calculating average revenue potential by room type
avg_revenue_by_room <- aggregate(revenue_potential ~ room_type, airbnb_data, mean)


barplot(
  avg_revenue_by_room$revenue_potential,
  names.arg = avg_revenue_by_room$room_type,
  col = "lightblue",
  main = "Average Revenue Potential by Room Type",
  xlab = "Room Type",
  ylab = "Average Revenue Potential"
)


```
##This bar chart shows the average revenue potential for different types of Airbnb rooms. Entire homes or apartments have the highest average revenue potential. Private rooms and shared rooms earn significantly less compared to entire homes or apartments.

```{r}
# Calculating total revenue potential by neighbourhood group
total_revenue_by_neighbourhood <- aggregate(revenue_potential ~ neighbourhood_group, airbnb_data, sum)

barplot(
  total_revenue_by_neighbourhood$revenue_potential,
  names.arg = total_revenue_by_neighbourhood$neighbourhood_group,
  col = "lightgreen",
  main = "Total Revenue Potential by Neighbourhood Group",
  xlab = "Neighbourhood Group",
  ylab = "Total Revenue Potential"
)


```
##This bar chart shows the total revenue potential for Airbnb listings in different New York City neighborhoods. Manhattan has the highest total revenue potential, followed by Brooklyn. Queens, the Bronx, and Staten Island have much lower total revenue potential in comparison.


Linear Regression to Predict Price

```{r}

model <- lm(price ~ number_of_reviews + availability_365 + minimum_nights, data = airbnb_data)

summary(model)

```

Check Model Performance
```{r}
# View the coefficients and R-squared value
print(summary(model)$coefficients)
print(paste("R-squared:", summary(model)$r.squared))

```
##The R-squared value (0.011) means the model explains only 1% of the changes in price, so the model is not very accurate.

##The coefficients show that more reviews slightly decrease the price, while more availability slightly increases the price.



 Making Predictions
```{r}

airbnb_data$predicted_price <- predict(model, airbnb_data)

head(airbnb_data[c("price", "predicted_price")])

```
 
 Plot Actual vs Predicted Prices
```{r}
plot(airbnb_data$price, airbnb_data$predicted_price,
     main = "Actual vs Predicted Prices",
     xlab = "Actual Price",
     ylab = "Predicted Price",
     col = "blue",
     pch = 16)

abline(0, 1, col = "red", lwd = 2)

```
##This scatter plot compares the actual prices of Airbnb listings to the predicted prices from the model. The blue dots represent individual listings, and the red line shows the ideal case where predicted prices match actual prices perfectly. Most of the predicted prices are far from the red line, which means the model does not predict the actual prices very accurately.
 